SIMD\slash SIMT-vectorized Thermo-chemical Source Term Evaluation

### 2018-01-31 derivations.tex
updates

@@

First, the derivative of the concentration of a species $k$ with respect to the state parameter is:
\begin{subequations}
 \label{e:dck_de}
 \begin{align}
 \lhs{dck}{\frac{\partial [C]_{k} }{\partial V }} &=
  \begin{dcases}
  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns  - 1} [C]_{k}} & \mathcond{dck}{k = \ns}
 \end{dcases}&\cond{dck}{\dconp} \\
 \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
 \begin{dcases}
   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns} \\
   \rhs{dck}{\frac{1}{T \Ru}} & \mathcond{dck}{k = \ns}
 \end{dcases}&\cond{dck}{\dconv}
 \end{align}
\end{subequations}
while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp~and \conv.
Following the now familiar collapsing of the specific-heat\slash concentration sum, and substituting in the energy source term gives:
\begin{subequations}
\begin{align}
 \frac{\partial\dot{T}}{\partial{V}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_p}_{k}} \Biggl[ - \sum_{k=1}^{-1 + \ns} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial V }_{k} + & \nonumber \\
				     & \mathindent[largerindentand] \frac{1}{V} \frac{\text{d} T }{\text{d} t } \sum_{k=1}^{-1 + \ns} \left(- {C_p}_{\ns} + {C_p}_{k}\right) [C]_{k}\Biggr] & \text{\dconp} \\
 \frac{\partial\dot{T}}{\partial{P}} &= \frac{1}{\sum_{k=1}^{\ns} [C]_{k} {C_v}_{k}} \Biggl[- \sum_{k=1}^{-1 + \ns} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \frac{\partial \dot{\omega} }{\partial P }_{k} - & \nonumber \\
				     & \mathindent[largerindentand] \frac{{C_v}_{\ns}}{T \Ru} \frac{\text{d} T }{\text{d} t } \Biggr] & \text{\dconv}
\end{align}
\end{subequations}
Finally, the species production rate derivative with respect to the state parameter will be computed in~\cref{s:d_spec_prod}.

@@

The derivative of the net molar rate of production of a species $k$ with respect to temperature is:
\begin{align}
 \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
 \intertext{or:}
   \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }
\end{align}

@@

Finally, the derivative of the net molar rate of production of a species $k$ with respect to the state parameter is:
\begin{align}
  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp}
  \intertext{and:}
  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv}
\end{align}

@@

The derivatives of the thermodynamic state parameter source terms with respect to temperature are given by differentiating~\cref{e:param_complete}:
\begin{align}
 \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial T} & \nonumber \\
		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(\frac{\partial \dot{\omega} }{\partial T }_{k} + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp}
\intertext{and:}
 \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial T} & \nonumber \\
		    &= \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(T \frac{\partial \dot{\omega} }{\partial T }_{k} + \dot{\omega}_{k}\right) & \nonumber \\
		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv}
\end{align}

@@

Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
\begin{align}
 \label{e:drop_dt}
 \frac{\partial R }{\partial T }_{i} =& \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right) - \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} \nonumber \\
				      &\qquad + \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}} - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}}
\end{align}
\subsubsection{Molar derivative}
\label{s:dri_dnj}
The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
\begin{equation}
 \label{e:dropf_dnj_1}
 \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_{f, i}} \;.
\end{equation}

@@

\subsubsection{Pressure-dependent reactions}
\label{s:dri_pdep}
The P-Log and Chebyshev pressure-dependent reactions, described by~\cref{e:kf_plog,e:kf_cheb} require separate treatment from the derivatives of strictly Arrhenius-based reaction rate formulations examined in~\cref{s:dri_dt,s:dri_dnj,s:dri_de}.
Beginning with P-Log reactions, the derivative of the forward rate coefficient with respect to temperature is:
\begin{align}
 \frac{\partial {k_{f, i}} }{\partial T } &= \biggl( \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \left(- \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{k_{2}} \frac{\text{d} k_2 }{\text{d} T }\right) \times  \Bigl( \nonumber \\
					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_{f, i}}
\end{align}
while the derivatives of $k_{1}$ and $k_{2}$ are given by~\cref{e:dkf_dt}---with the corresponding Arrhenius parameters (see~\cref{e:plog_k1,e:plog_k2}) substituted in.
Simplifying gives:
\begin{align}
 \label{e:dkf_plog_dt}
 \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru}\Biggr)
\end{align}
Similar to~\cref{s:dri_dt},~\cref{e:dkf_plog_dt} may be substituted into~\cref{e:dropfdt_1,e:droprdt_1} to give:
\begin{align}
 \frac{\partial R_{i} }{\partial T } &= \left(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \beta_1 + \frac{E_{a_1}}{T \Ru} \right) \times \nonumber \\
				     & \mathindent \left(\frac{{R_{f, i}} - {R_{r, i}}}{T}\right) + {R_{r, i}} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
\end{align}

@@

Next, the temperature derivative of Chevbyshev reactions will be considered:
\begin{align}
 \frac{\partial {k_{f, i}} }{\partial T } &= \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}
\end{align}
where $U_n$ is the Chebyshev polynomial of the second kind of degree n, expressed as:
\begin{equation}
 U_n \left(x\right) = \frac{\sin\left(\left(n + 1\right) \arccos x\right)}{\sin \left(\arccos x \right)}
\end{equation}
and:
\begin{equation}
 \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\min}} + \frac{1}{T_{\text{max}}}\right)}
\end{equation}
giving:
\begin{equation}
 \frac{\partial {k_{f, i}} }{\partial T } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)
\end{equation}
Following the same process for Arrhenius-based reactions (see~\cref{s:dri_dt}), the derivative of the net rate of progress with respect to temperature is:
\begin{align}
 \frac{\partial R }{\partial T }_{i} &= \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \left[ - \frac{2 \log{\left (10 \right )} \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right) \right] \left({R_{f, i}} - {R_{r, i}}\right) + \nonumber \\
				     & \mathindent \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} {R_{r, i}} + \frac{[C]}{T} \left(S^{\prime\prime}_{\ns} {k_{r, i}} - S^{\prime}_{\ns} {k_{f, i}}\right)
\end{align}

@@

>>>>

\label{e:dTdotdT_nojac}

@@

First, the derivative of the concentration of a species $k$ with respect to the state parameter is:
\begin{subequations}
 \label{e:dck_de}
 \begin{align}
 \lhs{dck}{\frac{\partial [C]_{k} }{\partial V }} &=
  \begin{dcases}
  \rhs{dck}{-\frac{[C]_{k}}{V}} & \mathcond{dck}{k \ne \ns} \\
  \rhs{dck}{\frac{1}{V} \sum_{k=1}^{\ns  - 1} [C]_{k}} & \mathcond{dck}{k = \ns}
 \end{dcases}&\cond{dck}{\dconp} \\
 \lhs{dck}{\frac{\partial [C]_{k} }{\partial P }} &=
 \begin{dcases}
   \rhs{dck}{\mathindent[s]0} & \mathcond{dck}{k \ne \ns} \\
   \rhs{dck}{\frac{1}{T \Ru}} & \mathcond{dck}{k = \ns}
 \end{dcases}&\cond{dck}{\dconv}
 \end{align}
\end{subequations}
while the specific heat and enthalpy\slash internal-energy are independent of the state parameter for both \conp~and \conv.
Following the now familiar collapsing of the specific-heat\slash concentration sum, and substituting in the energy source term gives:
\loadeq{dTdParam}
Finally, the species production rate derivative with respect to the state parameter will be computed in~\cref{s:d_spec_prod}.

@@

The derivative of the net molar rate of production of a species $k$ with respect to temperature is:
\begin{align}
 \label{e:dnkdotdT_compact}
 \mathcal{J}_{k + 2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial T } \nonumber \\
 \intertext{or:}
 \mathcal{J}_{k + 2, 1} & = V \frac{\partial \dot{\omega}_{k} }{\partial T }
\end{align}

@@

Finally, the derivative of the net molar rate of production of a species $k$ with respect to the state parameter is:
\begin{subequations}
\begin{align}
  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial V } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial V } & \nonumber \\
			 &= V \frac{\partial \dot{\omega}_{k} }{\partial V } + \dot{\omega}_{k} & \text{\dconp}
  \intertext{and:}
  \mathcal{J}_{k + 2, 2} &= \frac{\partial }{\partial P } \frac{\text{d} n_{k}}{\text{d} t} = \frac{\partial \dot{n}_k }{\partial P } & \nonumber \\
			 &=V \frac{\partial \dot{\omega}_{k} }{\partial P } & \text{\dconv}
\end{align}
\end{subequations}

@@

The derivatives of the thermodynamic state parameter source terms with respect to temperature are given by differentiating~\cref{e:param_complete}:
\begin{subequations}
\label{e:dEdotdT_nojac}
\begin{align}
 \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{V} }{\partial T} & \nonumber \\
		    &= \frac{V}{[C]} \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(\frac{\partial \dot{\omega} }{\partial T }_{k} + \frac{\dot{\omega}_{k}}{T}\right) \nonumber \\
		    &  \mathindent + \frac{V}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconp}
\intertext{and:}
 \mathcal{J}_{2, 1} &= \frac{\partial }{\partial T } \frac{\text{d} V}{\text{d} t} = \frac{\partial \dot{P} }{\partial T} & \nonumber \\
		    &= \Ru \sum_{k=1}^{-1 + \ns} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \left(T \frac{\partial \dot{\omega} }{\partial T }_{k} + \dot{\omega}_{k}\right) & \nonumber \\
		    & \mathindent + \frac{P}{T} \left(\frac{\text{d} \dot{T} }{\text{d} T } - \frac{1}{T} \frac{\text{d} T }{\text{d} t }\right) & \text{\dconv}
\end{align}
\end{subequations}

@@

Finally, combining~\cref{e:dropf_dt,e:dropr_dt} gives the total temperature derivative of the net rate of progress:
\loadeq{dropdT}
\subsubsection{Molar derivative}
\label{s:dri_dnj}
The derivative of the forward rate of progress with respect to the amount of moles of a species $j$ is:
\begin{equation}
 \label{e:dropf_dnj_1}
 \frac{d}{d n_{k}} {R_{f_i}} = \left(\frac{\partial}{\partial n_{j}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right) {k_{f, i}} \;.
\end{equation}

@@

\subsubsection{Pressure-dependent reactions}
\label{s:dri_pdep}
The P-Log and Chebyshev pressure-dependent reactions, described by~\cref{e:kf_plog,e:kf_cheb} require separate treatment from the derivatives of strictly Arrhenius-based reaction rate formulations examined in~\cref{s:dri_dt,s:dri_dnj,s:dri_de}.
Beginning with P-Log reactions, the derivative of the forward rate coefficient with respect to temperature is:
\begin{align}
 \frac{\partial {k_{f, i}} }{\partial T } &= \biggl( \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} \left(- \frac{1}{k_{1}} \frac{\text{d} k_1 }{\text{d} T } + \frac{1}{k_{2}} \frac{\text{d} k_2 }{\text{d} T }\right) \times  \Bigl( \nonumber \\
					 & \mathindent - \log{\left (P_{1} \right )} + \log{\left (P \right )}\Bigr)\biggr) {k_{f, i}}
\end{align}
while the derivatives of $k_{1}$ and $k_{2}$ are given by~\cref{e:dkf_dt}---with the corresponding Arrhenius parameters (see~\cref{e:plog_k1,e:plog_k2}) substituted in.
Simplifying gives:
\begin{align}
 \label{e:dkf_plog_dt}
 \frac{\partial {k_{f, i}} }{\partial T } &= \frac{{k_{f, i}}}{T} \Biggl(\frac{\left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right) \left(- \beta_1 + \beta_2 - \frac{E_{a_1}}{T \Ru} + \frac{E_{a_2}}{T \Ru}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \nonumber \\
					 & \mathindent[morespac] \beta_1 + \frac{E_{a_1}}{T \Ru}\Biggr)
\end{align}
Similar to~\cref{s:dri_dt},~\cref{e:dkf_plog_dt} may be substituted into~\cref{e:dropfdt_1,e:droprdt_1} to give:
\loadeq{dropplogdT}

@@

Next, the temperature derivative of Chevbyshev reactions will be considered:
\begin{align}
 \frac{\partial {k_{f, i}} }{\partial T } &= \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} \frac{\text{d} \tilde{T} }{\text{d} T } \left(j - 1\right) \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}
\end{align}
where $U_n$ is the Chebyshev polynomial of the second kind of degree n, expressed as:
\begin{equation}
 U_n \left(x\right) = \frac{\sin\left(\left(n + 1\right) \arccos x\right)}{\sin \left(\arccos x \right)}
\end{equation}
and:
\begin{equation}
 \frac{\text{d} \tilde{T}}{\text{d} T} = \frac{-2}{T^2 \left(-\frac{1}{T_{\min}} + \frac{1}{T_{\text{max}}}\right)}
\end{equation}
giving:
\begin{equation}
 \frac{\partial {k_{f, i}} }{\partial T } = \log{\left (10 \right )} {k_{f, i}} \sum_{\substack{1 \leq l \leq N_{P}\\1 \leq j \leq N_{T}}} - \frac{2 \mathcal{T}_{l - 1}\left(\tilde{P}\right) U_{j - 2}\left(\tilde{T}\right) \eta_{l,j}}{T^{2} \left(- \frac{1}{T_{\min}} + \frac{1}{T_{\max}}\right)} \left(j - 1\right)
\end{equation}
Following the same process for Arrhenius-based reactions (see~\cref{s:dri_dt}), the derivative of the net rate of progress with respect to temperature is:
\loadeq{dropchebdT}

@@

\section{Final jacobian form}
\label{s:jac_final}

Here we give a summary of the Jacobian derivations in the previous sections, as a complete reference guide to the equations evaluated by \texttt{pyJac}.

\subsection{Temperature derivatives}
In this section, the derivatives of all source terms with respect to the temperature are considered.

\subsubsection{Temperature source term derivatives}
Starting from~\cref{e:dTdotdT_nojac},~\cref{e:dnkdotdT_compact} is substituted in to yield:
\loadeq{dTdotdT}

\subsubsection{State parameter source term derivatives}
Similarly,~\cref{e:dnkdotdT_compact} is substituted into~\cref{e:dEdotdT_nojac}, giving:
\loadeq{dEdotdT}

\subsubsection{Molar source term derivatives}
From~\cref{e:dnkdotdT_compact} and~\cref{e:dwdot_dT}, a general form for the molar source term derivative with respect to temperature is obtained:
\loadeq{dnkdotdT}
As \texttt{pyJac} evaluates these derivative entries by looping over the reactions $i$ in the model, we convert~\cref{e:dnkdotdT} to an update form:
\begin{align}
 \lhs{dnkdotdT}{\mathcal{J}_{k + 2, 1}} &\pluseq  \rhs{dnkdotdT}{V \left(\nu_{k,i} R_{i} \frac{\partial c }{\partial T }_{i} + \nu_{k,i} \frac{\partial R }{\partial T }_{i} c_{i}\right)} & \mathcond{dnkdotdT}{k = 1, \ldots, \ns}
\end{align}
the rate of progress temperature derivative for all reaction types---excluding Chebyshev and P-Log---is:
\loadeq{dropdT}
while for P-Log reactions:
\loadeq{dropplogdT}
and for Chebyshev reactions:
\loadeq{dropchebdT}
